Relative Stable Maps in Genus One via Central Alignments

### 2018-04-29 genus-one-relative-maps.tex

\newcommand{\spec}{\operatorname{Spec}}
\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZco}[4]{\mathcal{V\!Z}^{\rm{ctr},\circ}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
\newcommand{\dvr}{\Delta}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\RR}{\textbf{R}}
\newcommand{\NN}{\operatorname{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

We now define the \emph{centrally alined relative space} to be the closure of the nice locus:
\begin{equation*} \VZc{1}{\alpha}{\PP^N|H}{d} := \overline{\VZco{1}{\alpha}{\PP^N|H}{d}} \subseteq \VZc{1}{n}{\PP^N}{d} \end{equation*}
This inherits the natural structure of a stack from the stack structure of the nice locus. It is proper since it is a closed substack of a proper stack, and it is irreducible because it is the closure of an irreducible stack. Hence it admits a fundamental class, which we can use to define \emph{reduced relative Gromov--Witten invariants}:
\begin{equation*} \langle \tau_{k_1}(\gamma_1), \ldots, \tau_{k_n}(\gamma_n) \rangle^{\PP^N|H, \operatorname{red}}_{0,\alpha,d} := \int_{[\VZc{1}{\alpha}{\PP^N|H}{d}]} \prod_{i=1}^n \ev_i^*(\gamma_i) \psi_i^{k_i} \end{equation*}

@@

\section{Recursion formula for $\PP^1|\infty, \Sigma \alpha = d$}
=======
\section{Recursion formula}

@@

\end{itemize}
\end{proof}

@@

\begin{lem}\label{Nice locus is irreducible}
The nice locus $\VZc{1}{\alpha}{\PP^N|H}{d}^{\circ}$ is irreducible.
\end{lem}
\begin{proof}
By definition the contraction radius $\delta$ is compatible with the map, i.e. the subcurve $C_0\subseteq C$ where $\lambda<\delta$ is equal to the maximal connected genus one subcurve contracted by $f$. Hence when the source curve is irreducible we must have $\delta=0$, and we see that the central alignment on $C$ is uniquely and trivially determined. Thus to specify a point in the nice locus we only need to specify the source curve $C$ (as a scheme) and the map $f$. A parametrisation can be given from the vector bundle:
\[ \operatorname{Vb}\left(\pi_*\OO_{\mathcal E}(\sum_{j=n+1}^{n+\delta}\sigma_j)\oplus\pi_*\OO_{\mathcal E}(\sum_{j=1}^{n+\delta}\sigma_j)^{\oplus r}\right) \quad \text{on} \quad \mathcal{M}_{1,n+\delta}\]
where $\pi\colon\mathcal E\to\mathcal M$ is the universal curve and $\delta=d-\sum\alpha_i$.
\end{proof}

>>>>

\newcommand{\spec}{\operatorname{Spec}}
\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZclb}[4]{\mathcal{V\!Z}^{\rm{ctr,lb}}_{#1,#2}(#3,#4)}
\newcommand{\VZco}[4]{\mathcal{V\!Z}^{\rm{ctr},\circ}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
\newcommand{\dvr}{\Delta}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\RR}{\textbf{R}}
\newcommand{\NN}{\operatorname{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

We now define the \emph{centrally aligned relative space} to be the closure of the nice locus:
\begin{equation*} \VZc{1}{\alpha}{\PP^N|H}{d} := \overline{\VZco{1}{\alpha}{\PP^N|H}{d}} \subseteq \VZc{1}{n}{\PP^N}{d} \end{equation*}
This inherits the natural structure of a stack from the stack structure of the nice locus. It is proper since it is a closed substack of a proper stack, and it is irreducible because it is the closure of an irreducible stack. Hence it admits a fundamental class, which we can use to define \emph{reduced relative Gromov--Witten invariants}:
\begin{equation*} \langle \tau_{k_1}(\gamma_1), \ldots, \tau_{k_n}(\gamma_n) \rangle^{\PP^N|H, \operatorname{red}}_{0,\alpha,d} := \int_{[\VZc{1}{\alpha}{\PP^N|H}{d}]} \prod_{i=1}^n \ev_i^*(\gamma_i) \cdot \psi_i^{k_i} \end{equation*}

@@

\section{Recursion formula for $\PP^1|\infty$, $\Sigma \alpha = d$.}
As in the genus zero case, we construct a line bundle (a quotient of jet bundles) and a section of that line bundle whose vanishing locus equals a union of boundary components (comb loci). (The moduli space for $\alpha+e_k$ does not appear because we are assuming maximal tangency.) These fit into the following families:
\subsection{$\mathcal{Y}^a$}
The first family is where the circuit is external:
\begin{equation*} \mathcal{Y}^a=\oM_{0,\lvert\alpha^{(0)}\rvert+r} \times\left(\VZc{1}{\alpha^{(1)}\cup\{m^{(1)}\}}{\PP^1|\infty}{d_1}\times_H {\prod_{i=2}^r} {}_H \M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^1|\infty}{d_i}\right) \end{equation*}

\subsection{$\mathcal{Y}^b$}
The second family is where the circuit is partly internal, partly external (and hence is a wheel of rational curves).
\begin{equation*} \mathcal Y^b=\oM_{0,\lvert\alpha^{(0)}\rvert+r}\times \left(\M{0}{\alpha^{(1)}\cup\{m^{(1)},m^{(2)}\}}{\PP^1|\infty}{d_1}\times_H \prod_{i=3}^r {}_H \M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^1|\infty}{d_i}\right)\end{equation*}

\subsection{$\mathcal{Y}^c$}
The third family is where the circuit is internal (and thus has zero degree, since we are working with $\PP^1$ relative $\infty$).
\begin{equation*} \mathcal Y^c = \oM^{\operatorname{ctr,lb}}_{1,\lvert\alpha^{(0)}\rvert+r} \times \prod_{i=1}^r {}_H \M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^1|\infty}{d_i} \end{equation*}
In the above, the first factor
\begin{equation*} \oM^{\operatorname{ctr,lb}}_{1,\lvert\alpha^{(0)}\rvert+r} \subseteq \oM^{\operatorname{ctr}}_{1,\lvert\alpha^{(0)}\rvert+r}  \end{equation*}
consists of those centrally aligned curves satisfying Gathmann's condition in $\Achow_0(C)$:
\begin{equation*} \sum_{x_j\in C}\alpha_jx_j-\sum_{i=1}^r m^{(i)}y_i= 0 \in \Achow_0(C)\end{equation*}
We saw before how to express this as a condition in $\Pic(E)$ where $E \subseteq C$ is the circuit. Notice that the condition above epends on the choice of $m^{(i)}$ but is really just a condition on the first factor of the product, so $\mathcal Y^c$ itself can be expressed as a product. \sidenote{Express the line bundle condition in terms of tautological classes?}

In this third case there is also a degeneracy condition that must be imposed on the second factor, which ensures that the map factors through the Smyth curve. To be more precise, the condition is that the derivatives
\begin{equation*} (\operatorname{d}\!f)|_{q_i} \in \operatorname{T}|_{\PP^1} \qquad i=1,\ldots,r \end{equation*}
are linearly dependent. Note that if $m_i > 1$ for any $i$, or if $r \geq 2$, then this is automatically satisfied. Thus the only case in which we have to impose this is for
\begin{equation*} \mathcal{Y}^c = \oM^{\operatorname{ctr,lb}}_{1,|\alpha^{(0)}|+r} \times {\M{0}{\alpha^{(i)}\cup (1}{\PP^1|\infty}{d}}^{\operatorname{degen}} \end{equation*}
where ``degen'' indicates that we impose degeneracy.

These spaces are all irreducible of the expected dimension, and it can be checked that they define divisors inside:
\begin{equation*} \VZc{1}{\alpha}{\PP^1|\infty}{d} \end{equation*}
In fact, this applies to all of the above spaces \emph{except} for the special case mentioned above (where we impose degeneracy). This has too small dimension, and hence will not appear in the recursion formula.

\section{Recursion formula}

@@

 \end{itemize}

\end{proof}

@@

\begin{lem}\label{Nice locus is irreducible}
The nice locus $\VZc{1}{\alpha}{\PP^N|H}{d}^{\circ}$ is irreducible.
\end{lem}
\begin{proof}
By definition the contraction radius $\delta$ is compatible with the map, i.e. the subcurve $C_0\subseteq C$ where $\lambda<\delta$ is equal to the maximal connected genus one subcurve contracted by $f$. Hence when the source curve is irreducible we must have $\delta=0$, and we see that the central alignment on $C$ is uniquely and trivially determined. Thus to specify a point in the nice locus we only need to specify the source curve $C$ (as a scheme) and the map $f$. A parametrisation can be given from the vector bundle:
\[ \operatorname{Vb}\left(\pi_*\OO_{\mathcal E}(\sum_{j=n+1}^{n+\delta}\sigma_j)\oplus\pi_*\OO_{\mathcal E}(\sum_{j=1}^{n+\delta}\sigma_j)^{\oplus r}\right) \quad \text{on} \quad \mathcal{M}_{1,n+\delta}\]
where $\pi\colon\mathcal E\to\mathcal M$ is the universal curve and $\delta=d-\sum\alpha_i$. Notice that this is a vector bundle because the source curve is smooth of genus one.
\end{proof}
